# Fonts in R

## Or, _Trying to spruce up your graphics_

#### Some generic tips and pointers

[An example](http://stackoverflow.com/questions/30058107/using-custom-otf-fonts-in-ggplot2) 

Note that Myriad Pro and Garamond Pro are both OTF type fonts--not TTF. This means that the `showtext` package is our solution, since `extrafonts` does not (at this writing) seem to support OTF fonts.


---

### `showtext`

[CRAN page: showtext: Using Fonts More Easily in R Graphs](https://cran.rstudio.com/web/packages/showtext/index.html)

[package manual](https://cran.rstudio.com/web/packages/showtext/showtext.pdf)

[Using system fonts in R graphs](http://statr.me/2014/01/using-system-fonts-in-r-graphs/)

[showtext: Using System Fonts in R Graphics](https://journal.r-project.org/archive/2015-1/qiu.pdf) [pdf file]

[`showtext` page on github](https://github.com/yixuan/showtext)



```{r}

# an example chart with default fonts

library(ggplot2)

fonttestplot <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Miles Per Gallon by Weight of Car")

fonttestplot

# install `showtext`
install.packages("showtext")

library(showtext)

# add MyriadPro as font
font.paths()
font.files()


#font.add("MyriadPro", "C:\\Windows\\Fonts\\MyriadPro-Regular.otf")
font.add("MyriadPro", "MyriadPro-Regular.otf")

fonts()

fonttestplot <- fonttestplot + theme(text=element_text(family="Times"))

showtext.auto()

fonttestplot

# run chart again, with Myriad Pro font


```




---

### `extrafont`

[Tired of using Helvetica in your R graphics? Here’s how to use the fonts you like](http://zevross.com/blog/2014/07/30/tired-of-using-helvetica-in-your-r-graphics-heres-how-to-use-the-fonts-you-like-2/)

[How to use your favorite fonts in R charts](http://blog.revolutionanalytics.com/2012/09/how-to-use-your-favorite-fonts-in-r-charts.html)

* from 2012; says support for OTF fonts is experimental and you need the development version of the package to get it to work

[Change fonts in ggplot2, and create xkcd style graphs](http://www.r-bloggers.com/change-fonts-in-ggplot2-and-create-xkcd-style-graphs/)


#### Process for getting `extrafont` to work on your computer

```{r}

install.packages("extrafont")

library(extrafont)
font_import()


```





