# Fonts in R

## Or, _Trying to spruce up your graphics_

#### Some generic tips and pointers

[An example](http://stackoverflow.com/questions/30058107/using-custom-otf-fonts-in-ggplot2) 

Note that Myriad Pro and Garamond Pro are both OTF type fonts--not TTF. This means that the `showtext` package is our solution, since `extrafonts` does not (at this writing) seem to support OTF fonts.


---

### `showtext`

[CRAN page: showtext: Using Fonts More Easily in R Graphs](https://cran.rstudio.com/web/packages/showtext/index.html)

[package manual](https://cran.rstudio.com/web/packages/showtext/showtext.pdf)

[`showtext` page on github](https://github.com/yixuan/showtext)

[R Journal article](https://journal.r-project.org/archive/2015-1/qiu.pdf) [PDF file]

[Using system fonts in R graphs](http://statr.me/2014/01/using-system-fonts-in-r-graphs/)

[showtext: Using System Fonts in R Graphics](https://journal.r-project.org/archive/2015-1/qiu.pdf) [pdf file]

[Using system and web fonts in R plots](http://www.magesblog.com/2015/06/using-system-and-web-fonts-in-r-plots.html)

[another stackoverflow thing on fonts](http://stackoverflow.com/questions/1395323/fonts-in-r-plots)

```{r}

setwd("C:/@data/GitHub/WorkInProgress")

# an example chart with default fonts

library(ggplot2)

fonttestplot <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Miles Per Gallon by Weight of Car")

fonttestplot

```




```{r}


download.file("http://fontpro.com/download-family.php?file=36289",
              "newrocker.ttf", mode="wb")
download.file("http://fontpro.com/download-family.php?file=35128",
              "cutetumblr.ttf", mode ="wb")

library(showtext)

font.add("newrocker", "newrocker.ttf")
font.add("cutetumblr", "cutetumblr.ttf")

png("showtext-ex2.png", 800, 500)
plot(1, type = "n")
showtext.begin()
text(1, 1.2, "Let me tell you a story", cex = 4, family = "newrocker")
text(1, 0.8, "Long long ago...", cex = 4, family = "cutetumblr")
showtext.end()
dev.off()


font.add("Garamond", "GARA.TTF")
font.add("MyriadPro", "MyriadPro-Regular.otf")

png("showtext-ex2a.png", 800, 500)
plot(1, type = "n")
showtext.begin()
text(1, 1.2, "Let me tell you a story", cex = 4, family = "Garamond")
text(1, 0.8, "Long long ago...", cex = 4, family = "MyriadPro")
showtext.end()
dev.off()



```

#### Ok, now we are getting somewhere! 

These chunks seem to work ... at producing a PNG file.

```{r}
# install `showtext`
install.packages("showtext")

library(showtext)

#showtext.auto()


# add MyriadPro as font
font.paths()
font.files()

#font.add("MyriadPro", "C:\\Windows\\Fonts\\MyriadPro-Regular.otf")
font.add("Garamond", "GARA.TTF")
font.add("MyriadPro", "MyriadPro-Regular.otf")


# plot inside the png() etc instructions

library(ggplot2)

png("showtext-cars.png", 800, 500)
plot(1, type = "n")
showtext.begin()

fonttestplot <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Miles Per Gallon by Weight of Car") +
  theme(plot.title = element_text(family="Garamond"),
        axis.title.x = element_text(family="MyriadPro"),
        axis.title.y = element_text(family="MyriadPro"))
fonttestplot

showtext.end()
dev.off()


```

```{r}

# create plot first, render to Plot window


library(ggplot2)
library(showtext)

fonttestplot2 <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Miles Per Gallon by Weight of Car") +
  theme(plot.title = element_text(family="Garamond"),
        axis.title.x = element_text(family="MyriadPro"),
        axis.title.y = element_text(family="MyriadPro"))

fonttestplot2

# then create png of existing object

png("showtext-cars2.png", 800, 500)
plot(1, type = "n")
showtext.begin()

fonttestplot2

showtext.end()
dev.off()


```


### A plot in a Windows graphic window

```{r}


download.file("http://fontpro.com/download-family.php?file=36289",
              "newrocker.ttf", mode="wb")
download.file("http://fontpro.com/download-family.php?file=35128",
              "cutetumblr.ttf", mode ="wb")

library(showtext)

font.add("newrocker", "newrocker.ttf")
font.add("cutetumblr", "cutetumblr.ttf")

windows(800, 500)
plot(1, type = "n")
showtext.begin()
text(1, 1.2, "Let me tell you a story", cex = 4, family = "newrocker")
text(1, 0.8, "Long long ago...", cex = 4, family = "cutetumblr")
showtext.end()

# and then when you're ready to close the window...
dev.off()


```

# SUMMARY CODE - THIS IS ALL YOU NEED

The `fonttestplot2` object, rendered in a Windows window and then as a .png file

```{r}

# load packages
library(ggplot2)
library(showtext)

# add MyriadPro as font
font.paths()
font.files()

#font.add("MyriadPro", "C:\\Windows\\Fonts\\MyriadPro-Regular.otf")
font.add("Garamond", "GARA.TTF")
font.add("MyriadPro", "MyriadPro-Regular.otf")


# ggplot2 code to create plot object
fonttestplot2 <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Miles Per Gallon by Weight of Car") +
  theme(plot.title = element_text(family="Garamond"),
        axis.title.x = element_text(family="MyriadPro"),
        axis.title.y = element_text(family="MyriadPro"))


# ---

# render in a Windows window

windows(800, 500)
plot(1, type = "n")
showtext.begin()

fonttestplot2

showtext.end()

# and then when you're ready to close the window...
dev.off()


# ---

# then save a png of existing object

png("showtext-cars2.png", 800, 500)
plot(1, type = "n")
showtext.begin()

fonttestplot2

showtext.end()
dev.off()


```



## DO NOT CROSS THIS LINE

---

```{r}

library(showtext)
## Loading Google fonts (http://www.google.com/fonts)
font.add.google("Gochi Hand", "gochi")
font.add.google("Schoolbell", "bell")
font.add.google("Covered By Your Grace", "grace")
font.add.google("Rock Salt", "rock")

## Automatically use showtext to render text for future devices
showtext.auto()

## Tell showtext the resolution of the device,
## only needed for bitmap graphics. Default is 96
## showtext.opts(dpi = 96)

set.seed(123)
x = rnorm(10)
y = 1 + x + rnorm(10, sd = 0.2)
y[1] = 5
mod = lm(y ~ x)

## Plotting functions as usual
## Open a graphics device if you want, e.g.
## png("demo.png", 700, 600, res = 96)

op = par(cex.lab = 2, cex.axis = 1.5, cex.main = 2)
plot(x, y, pch = 16, col = "steelblue",
     xlab = "X variable", ylab = "Y variable", family = "gochi")
grid()
title("Draw Plots Before You Fit A Regression", family = "bell")
text(-0.5, 4.5, "This is the outlier", cex = 2, col = "steelblue",
     family = "grace")
abline(coef(mod))
abline(1, 1, col = "red")
par(family = "rock")
text(1, 1, expression(paste("True model: ", y == x + 1)),
     cex = 1.5, col = "red", srt = 20)
text(0, 2, expression(paste("OLS: ", hat(y) == 0.79 * x + 1.49)),
     cex = 1.5, srt = 15)
legend("topright", legend = c("Truth", "OLS"), col = c("red", "black"), lty = 1)

par(op)


# ------------
# IN A SEPARATE WINDOWS WINDOW

windows(800, 500)
#plot(1, type = "n")
showtext.begin()

set.seed(123)
x = rnorm(10)
y = 1 + x + rnorm(10, sd = 0.2)
y[1] = 5
mod = lm(y ~ x)

## code from above to render the chart

op = par(cex.lab = 2, cex.axis = 1.5, cex.main = 2)
plot(x, y, pch = 16, col = "steelblue",
     xlab = "X variable", ylab = "Y variable", family = "gochi")
grid()
title("Draw Plots Before You Fit A Regression", family = "bell")
text(-0.5, 4.5, "This is the outlier", cex = 2, col = "steelblue",
     family = "grace")
abline(coef(mod))
abline(1, 1, col = "red")
par(family = "rock")
text(1, 1, expression(paste("True model: ", y == x + 1)),
     cex = 1.5, col = "red", srt = 20)
text(0, 2, expression(paste("OLS: ", hat(y) == 0.79 * x + 1.49)),
     cex = 1.5, srt = 15)
legend("topright", legend = c("Truth", "OLS"), col = c("red", "black"), lty = 1)

par(op)

showtext.end()

dev.off()


```


```{r}
# copy-paste from 
# https://github.com/yixuan/showtext

library(showtext)

## Add fonts that are available on Windows
font.add("heiti", "simhei.ttf")
font.add("constan", "constan.ttf", italic = "constani.ttf")


library(ggplot2)

p = ggplot(NULL, aes(x = 1, y = 1)) + ylim(0.8, 1.2) +
    theme(axis.title = element_blank(), axis.ticks = element_blank(),
          axis.text = element_blank()) +
    annotate("text", 1, 1.1, family = "heiti", size = 15,
             label = "\u4F60\u597D\uFF0C\u4E16\u754C") +
    annotate("text", 1, 0.9, label = 'Chinese for "Hello, world!"',
             family = "constan", fontface = "italic", size = 12)

showtext.auto()  ## automatically use showtext for new devices

print(p)  ## on-screen device

showtext.auto(FALSE)  ## turn off if no longer needed

# so that works...

# tweaks

#font.add("MyriadPro", "C:\\Windows\\Fonts\\MyriadPro-Regular.otf")
font.add("garamond", "GARA.TTF")
font.add("myriadpro", "MyriadPro-Regular.otf")

p1 = ggplot(NULL, aes(x = 1, y = 1)) + ylim(0.8, 1.2) +
    theme(axis.title = element_blank(), axis.ticks = element_blank(),
          axis.text = element_blank()) +
#    annotate("text", 1, 1.1, family = "heiti", size = 15,
#             label = "\u4F60\u597D\uFF0C\u4E16\u754C") +
    annotate("text", 1, 0.9, label = 'Chinese for "Hello, world!"',
             family = "garamond", fontface = "italic", size = 12)

showtext.auto()  ## automatically use showtext for new devices

print(p1)  ## on-screen device


```




---

### `extrafont`

[Tired of using Helvetica in your R graphics? Here’s how to use the fonts you like](http://zevross.com/blog/2014/07/30/tired-of-using-helvetica-in-your-r-graphics-heres-how-to-use-the-fonts-you-like-2/)

[How to use your favorite fonts in R charts](http://blog.revolutionanalytics.com/2012/09/how-to-use-your-favorite-fonts-in-r-charts.html)

* from 2012; says support for OTF fonts is experimental and you need the development version of the package to get it to work

[Change fonts in ggplot2, and create xkcd style graphs](http://www.r-bloggers.com/change-fonts-in-ggplot2-and-create-xkcd-style-graphs/)


#### Process for getting `extrafont` to work on your computer

```{r}

install.packages("extrafont")

library(extrafont)
font_import()


```


#### Devices

Understanding what devices are in use

* [dev {grDevices}: Control Multiple Devices](https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/dev.html)



-30-
