---
title: "ggplot2_palette eigengrau"
author: "Martin Monkman"
date: "March 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If you are an R user creating data visualizations (plots, charts, graphs, and maps), you will be aware that there are _many_ colours available to you. Much is available in base R!

```{r}
library(grDevices)
library(RColorBrewer)

colors <- brewer.pal(10, "BuPu")
palette(colors)

pie_data_10 <- c(10, 10, 10, 10, 10, 10, 10, 10, 10, 10)

pie(pie_data_10)

```


## ggplot2 palettes

The R package `ggplot2` provides access to a variety of built-in palettes, sometimes you just need more.

* `RColorBrewer` - this package is based on the work of cartographer [Cynthia Brewer](http://www.personal.psu.edu/cab38/)
   + note that there are three types of palette: sequential, diverging, and qualitative.
   + the color schemes can be found at [colorbrewer2.org](http://colorbrewer2.org/)

```{r}

display.brewer.all()

```

   
## eigengrau palette

You can also create your own custom palettes. Here's the code to create a custom palette based on the not-quite-black colour we see in the pitch black darkness, often refered to as "eigengrau". In hex, this colour is [#16161D](http://www.color-hex.com/color/16161d).

```{r}
# 4 step eigengrau palette
eigengrau_palette <- c("#16161D", "#5b5b60", "#a1a1a4", "#e7e7e8")

# 2 step eigengrau palette (from eigengrau to palest)
eigengrau_palette <- c("#16161D", "#e7e7e8")

pal <- colorRampPalette(eigengrau_palette)
pal10 <- pal(10)

pie_data_10 <- c(10, 10, 10, 10, 10, 10, 10, 10, 10, 10)

pie(pie_data_10, col = pal10)

# repeat with 5 shades

eigengrau_palette <- c("#16161D", "#e7e7e8")

pal <- colorRampPalette(eigengrau_palette)
pal05 <- pal(5)

pie_data_05 <- c(5, 5, 5, 5, 5)
pie_data_05 <- c(1, 2, 3, 4, 5)

pie(pie_data_05, col = pal05)


```




### Further reading

* ["Colors in R"](http://simplystatistics.org/2011/10/17/colors-in-r/), from [Simply Statistics](http://simplystatistics.org/)

* Nathan Yau, ["Using Color Scales and Palettes in R"](http://thedata.co/sites/thedata.co/files/u1/VisThis_Tutorial_Using%20Color%20Scales%20and%20Palettes%20in%20R.pdf) {PDF}, at [thedata.co](thedata.co)


-30-