---
title: "goal scoring"
output: html_notebook
---


 
 
 <blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">imagine a hockey game where we know that Team A scores exactly 1 goal for sure and Team B takes 20 shots, each with a 5.5% chance of going in.<br>which team would you rather be?<br>(nothing additional happens if you tie.)</p>&mdash; Namita (@nnstats) <a href="https://twitter.com/nnstats/status/1075189243314610176?ref_src=twsrc%5Etfw">December 19, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


response blog post: https://hockey-graphs.com/2018/12/19/some-people-were-wrong-on-twitter/


<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">x &lt;- replicate(10000, sum(sample(0:1, 20, TRUE, c(0.945, 0.055))))<br>table(ifelse(x == 0, &quot;Team A win&quot;, ifelse(x == 1, &quot;Draw&quot;, &quot;Team B win&quot;))) / 100<br><br>Draw Team A win Team B win <br>37.52      32.18      30.30<a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/Kv8rXj7AIi">https://t.co/Kv8rXj7AIi</a></p>&mdash; Eric (@expersso) <a href="https://twitter.com/expersso/status/1075425581146271747?ref_src=twsrc%5Etfw">December 19, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

```{r}

set.seed(894) #number of regular season NHL goals Wayne Gretzky scored 
 
x <- replicate(10000, sum(sample(0:1, 20, TRUE, c(0.945, 0.055))))

table(ifelse(x == 0, "Team A win", ifelse(x == 1, "Draw", "Team B win"))) / 100

```


### tally the number of games by the number of goals scored

```{r}

library(tidyverse)

x <- as.tibble(x)

x %>%
  group_by(value) %>%
  tally()

```


a plot of the above table

```{r}

goalsscored_plot <-
ggplot(data=x, aes(value)) + 
  geom_histogram(binwidth = 1, fill = "white", colour = "black") +
  scale_x_continuous(breaks=seq(0,8,1)) +
  theme(panel.grid.major.x = element_blank()) +
  labs(x = "goals per 20 shots",
       y = "number of games (out of 10,000",
       title = "Distribution of games by number of goals scored",
       subtitle = "20 shots per game each with 5.5% chance of scoring",
       caption = "source: a question @nnstats posed on twitter")


ggsave(filename = "goalsscored_plot.png", plot = goalsscored_plot,
       width = 8, height = 6)

```



-30-