---
title: "goal scoring"
output: html_notebook
---


 
 
 <blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">imagine a hockey game where we know that Team A scores exactly 1 goal for sure and Team B takes 20 shots, each with a 5.5% chance of going in.<br>which team would you rather be?<br>(nothing additional happens if you tie.)</p>&mdash; Namita (@nnstats) <a href="https://twitter.com/nnstats/status/1075189243314610176?ref_src=twsrc%5Etfw">December 19, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>



<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">x &lt;- replicate(10000, sum(sample(0:1, 20, TRUE, c(0.945, 0.055))))<br>table(ifelse(x == 0, &quot;Team A win&quot;, ifelse(x == 1, &quot;Draw&quot;, &quot;Team B win&quot;))) / 100<br><br>Draw Team A win Team B win <br>37.52      32.18      30.30<a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/Kv8rXj7AIi">https://t.co/Kv8rXj7AIi</a></p>&mdash; Eric (@expersso) <a href="https://twitter.com/expersso/status/1075425581146271747?ref_src=twsrc%5Etfw">December 19, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

```{r}
 
x <- replicate(10000, sum(sample(0:1, 20, TRUE, c(0.945, 0.055))))

table(ifelse(x == 0, "Team A win", ifelse(x == 1, "Draw", "Team B win"))) / 100

```


### make a histogram of the distribution

```{r}

library(tidyverse)

x <- as.tibble(x)

x %>%
  group_by(value) %>%
  tally()

ggplot(data=x, aes(value)) + 
  geom_histogram(binwidth = 1, fill = "white", colour = "black")


```



-30-