---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Using `purrr` and `broom` in conjunction.

Some background reading:

`broom`

* https://cran.r-project.org/web/packages/broom/vignettes/broom.html

* https://github.com/tidyverse/broom/issues/158




## preliminaries

packages

```{r}

library(tidyverse)
library(broom)
library(purrr)

library(skimr)

library(beepr)

beepr::beep("coin")


```


## data - build a table

```{r}

set.seed(8675309)

bino1 = rbinom(500, 1, 0.5)

data_for_mod <- as.data.frame(bino1)

data_for_mod <- data_for_mod %>%
  mutate(bino2 = as.factor(if_else(bino1 == 1, rbinom(500, 1, 0.75),
                                  rbinom(500, 1, 0.1))))
           
data_for_mod <- data_for_mod %>%
  mutate(bino3 = as.factor(if_else(bino1 == 1, rbinom(500, 4, 0.75),
                                  rbinom(500, 4, 0.1))))


data_for_mod <- data_for_mod %>%
  mutate(con1 = if_else(bino1 == 1, round(rnorm(500, mean = 50, sd = 10), 3),
                                  round(rnorm(500, mean = 25, sd = 5), 3))
  )

data_for_mod <- data_for_mod %>%
  mutate(con2 = round(runif(500, min = 0, max = 1), 3))



data_for_mod

```


### Testing `skimr`

See https://ropensci.org/blog/2017/07/11/skimr/ 

(alas, earwax: Windows does not render the histogram / sparklines correctly)

```{r}

skim(data_for_mod)

```


## purrr the lm

modified from https://sebastiansauer.github.io/multiple-lm-purrr2/


```{r}

data_for_mod %>% 
  dplyr::select(-bino1) %>%  # exclude outcome, leave only predictors 
  dplyr::select(-c(bino2, bino3)) %>%  # exclude other factors, leave only continuous (double) variables 
  map(~lm(data_for_mod$bino1 ~ .x, data = data_for_mod)) %>% 
  map(summary) %>% 
  map_dbl("r.squared") %>% 
  tidy %>% 
  dplyr::arrange(desc(x)) %>% 
  rename(r.squared = x) -> r2s

print(r2s)

```




### glm version

#### basic

```{r}

mod1 = glm(bino1 ~ bino2, family = binomial, data_for_mod)

summary(mod1)

tidy(mod1)

mod1 %>%
  tidy() %>%
  filter(term == "bino21") %>%
  pull("estimate")


```



* modify the first part of the above code to select only the binomial (factor) variables
* modify the model type to `glm`

It works!



### make it purrr

This chunk produces the "summary" output for the individual models for both X variables

```{r}

data_for_mod %>% 
  dplyr::select(-bino1) %>%  # exclude outcome, leave only predictors 
  dplyr::select(-c(con1, con2)) %>%  # exclude other factors, leave only continuous (double) variables 
  map(~glm(data_for_mod$bino1 ~ .x, data = data_for_mod, family = binomial))

```



...but with the `tidy()` etc rows from above (with the term name changed) it fails
 -- error "Error in tidy.list(.) : No tidying method recognized for this list"

Note: I have tried a variety of filter options!

```{r}
glm_list <- data_for_mod %>% 
  dplyr::select(-bino1) %>%  # exclude outcome, leave only predictors 
  dplyr::select(-c(con1, con2)) %>%  # exclude other factors, leave only continuous (double) variables 
  map(~glm(data_for_mod$bino1 ~ .x, data = data_for_mod, family = binomial))
```

** NOTE: what follows is from @cwickham **

The problem is, at this stage we have a list of `glm` objects:
```{r}
glm_list
```

If we try to pipe that into `tidy()` we get the error because `tidy()` sees it as a list, (not a glm object):
```{r}
glm_list %>% tidy()
```

Really you want to `map()` `tidy()` (i.e. for each mdoel in this list, tidy it):
```{r}
glm_list %>% map(tidy)
```

Then you could proceed to pull out the pieces you want, but you would have to keep mapping because we still have a list of data frames:
```{r}
glm_list %>% 
  map(tidy) %>%
  map("estimate")
```
