---
title: "2 hours"
author: "Martin Monkman"
date: "April 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

From @StatsbyLopez:
> David and Emily arrive at the Rec Centre at times uniformly distributed between noon and 2 pm independently of each other. What's the probability that they will get there within 20 minutes of each other?
[https://twitter.com/StatsbyLopez/status/852898752876642305](https://twitter.com/StatsbyLopez/status/852898752876642305)


My solution: 

build a data frame with two variables, each with a value between 1 and 120 (the number of minutes in a 2 hour period) randomly selected from a uniform distibution distribution

plot as X-Y

then mutate to calculate difference between the two 

calculate proportion where difference <= 20


```{r}

library(dplyr)
library(tibble)
library(ggplot2)

```


```{r}

set.seed(8675309)

# with a thousand

ETA <- data_frame(Emily = floor(runif(1000, 1, 120)),
                  David = floor(runif(1000, 1, 120))
)

ggplot(ETA, aes(x = Emily, y = David)) +
  geom_point()


# with a million cases

ETA <- data_frame(Emily = floor(runif(1e6, 1, 120)),
                  David = floor(runif(1e6, 1, 120))
)

ETA <- ETA %>%
  mutate(difference = Emily - David)


# plot a histogram of the distribution
ggplot(ETA, aes(x = difference)) +
  geom_histogram(binwidth = 10) +
  geom_density()


# summary table of the 
ETA %>%
  mutate(within20 = ifelse((difference >= -20 & difference <= 20), "Yes", "No")) %>%
  group_by(within20) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
  




```


### David Robinson's elegant solution

[https://twitter.com/drob/status/852899563782451200](https://twitter.com/drob/status/852899563782451200)


```{r}

mean(abs(runif(1e7, 0, 2) - runif(1e7, 0, 2)) < 1 / 3)

```

