---
title: "multiple evaluation"
author: "Martin Monkman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## The problem

I have a "check all that apply" survey question that has been coded with as many columns as there are options, but the columns are only filled if a respondent has checked an option.

I want to create a new variable for each option, which has a TRUE or FALSE for the options.

Here's a toy example:

```{r data}

dat <- tribble(
  ~user_id, ~var_01, ~var_02, ~var_03,
  "user_w", "option 1", "option 2", "option 3",
  "user_x", "option 1", "option 2", NA,
  "user_y", "option 2", "option 3", NA,
  "user_z", "option 3", NA, NA
)

dat

```

I could use a `case_when()` inside a `mutate()`, but it has to be repeated for as many options as I have. (IRL one of the surveys I have been asked to analyze has 14 options...so it would require 14 evaluations for each of the 14 mutations. That's 196 copy/paste jobs.) 

Here's the code for option 2, which occurs in var_01 and var_02.

```{r}

dat %>% 
  mutate(option_2 = case_when(
    var_01 == "option 2" ~ TRUE,
    var_02 == "option 2" ~ TRUE,
    var_03 == "option 2" ~ TRUE,
    TRUE ~ FALSE
  ))

```

What I have been trying to find is a way to use something like `across()` which will evaluate all `var_` variables in a single mutate line. 

* All of the `across()` examples I have found apply the same function repeatedly to all the variables, e.g. mean of var_01, mean of var_02, etc.

* If the data was stored in a single variable (i.e. a "pick one" survey question, rather than a "check all that apply") a `pivot_wider()` with a TRUE / FALSE option

I have a nagging sense that I am overlooking something that would give me a streamlined solution to this (other than a time machine, and politely asking whoever coded the data to use the option = T/F approach!)


## not quite solutions

https://stackoverflow.com/questions/58686186/recode-multiple-polytomous-variables

https://stackoverflow.com/questions/43836503/recoding-multiple-variables-in-r

dplyr::across() https://dplyr.tidyverse.org/articles/colwise.html



https://www.r-bloggers.com/2022/08/analyzing-multiple-response-questions/

{expss} 

* https://cran.r-project.org/web/packages/expss/index.html

* https://github.com/gdemin/expss


Note: the {expss} package has a set of functions to calculate summary tables from data that is stored in this way (_polytomous_). Here the `cross_cases()` function produces the frequency counts for our data.


```{r}
library(expss)
expss::cross_cases(dat, mrset(var_01 %to% var_03), total(label = "freq"))
```

Similarly, the `cross_cpct()` function calculates the column percentages:

```{r}
cross_cpct(dat, mrset(var_01 %to% var_03), total(label = "pct"))
```

But none of these options provide a way to mutate a new column into my original records.

