---
title: "split_strings"
author: "Martin Monkman"
date: "05/05/2021"
output: html_document
---


```{r setup}
library(tidyverse)
```


Some messy data--note that the 2nd row has an explicit missing Sales value in the middle and the end (there is a comma at the end of the string), and the 3rd row has an implicit missing Sales value at the end.

```{r}
df <- tribble(
  ~"ID", ~"City", ~"Sales",
  1, "Istanbul,Tokyo,Brussels", "4,5,6",
  2, "London,New York,Paris,Berlin,Madrid", "20,30,,50,",
  3, "London,Paris,New York,Munich", "1979,1980,1981"
)

df
```

The goal is to split each City cell into separate entry, and then match with the Sales column. Once that's done for each row, they get bound together into a single table.

The first row will look like this:

```{r}

tribble(
  ~"City", ~"Sales",
  "Istanbul", "4",
  "Tokyo", "5",
  "Brussels", "6"
)


```


Some experimentation

```{r}

df[1, , drop = TRUE]

df[1,]


df2 <- as.matrix(df)
as.vector(df2[1,])


df %>% 
  select(-ID) %>% 
  as.matrix() 

```


Got it! split the columns into separate objects, then use `data.frame()` to join them by position, then `rbind()`

```{r}
city_names <- str_split(df$City, ",")
sales <- str_split(df$Sales, ",")

df_a <- data.frame(city_names[[1]], sales[[1]])
colnames(df_a) <- c("city_name", "sales") 
  

df_b <- data.frame(city_names[[2]], sales[[2]])
colnames(df_b) <- c("city_name", "sales") 

rbind(df_a, df_b)

```

Turn it into a function

```{r}

city_names <- str_split(df$City, ",")
sales <- str_split(df$Sales, ",")

temp_df <- data.frame(city_names = vector("character"),
                     sales = vector("character"))


for(i in nrow(df)) {
  temp_df[i] <- data.frame(city_names[[i]], sales[[i]])
}

```

