---
title: "What is your VW New Beetle worth?"
author: "Martin Monkman"
date: "Monday, March 09, 2015"
output: html_document
---

Inputs:  current asking price, age, and KM of VW New Beetles available for sale throughout British Columbia,
as advertised on

* autotrader.ca 

* usedvictoria.ca

* craigslist.ca

The first chunk creates two linear regression models and plots them: 

## predicting price by KM


```{r}
library("ggplot2")

beetles <- read.csv("beetles.csv")

lm(formula = Price ~ KM, data = beetles)

plot_PKM <- ggplot(beetles, aes(x=KM, y=Price))
plot_PKM + geom_point() + stat_smooth(method=lm, level=0.95)

```

## predicting price by age

```{r}

lm(formula = Price ~ age, data = beetles)

plot_Page <- ggplot(beetles, aes(x=age, y=Price))
plot_Page + geom_point() + stat_smooth(method=lm, level=0.95)


plot_Page <- ggplot(beetles, aes(x=age, y=Price))
plot_Page + 
  geom_point() + 
  stat_smooth(method=lm, level=0.95) +
  theme_bw() +
  theme(legend.position="bottom")

```

## predicting price by KM and age

Now, a multiple regression model predicting price with both KM and age as the independent variables.

```{r}

lm(formula = Price ~ age + KM, data = beetles)

```


_How to plot this puppy?_


-30-