---
title: "Stephanie's fence"
author: "Martin Monkman"
date: "February 27, 2019"
output:
  github_document:
    html_preview: true
---

<!-- 
This file is licensed with the Creative Commons Attribution 4.0 International License.
-->

Note: https://github.com/bcgov/envreportutils

## Stephanie's fencepost


```{r}
library(weathercan)
library(dplyr)
library(ggplot2)

library(feather)

```


(1) Get and wrangle the [Environment and Climate Change Canada (ECCC)](http://climate.weather.gc.ca/historical_data/search_historic_data_e.html) [wind direction data](http://climate.weather.gc.ca/glossary_e.html#windSpd) using the the [rOpenSci](https://ropensci.org/) [`weathercan` R package](http://ropensci.github.io/weathercan/) `r emo::ji("package")`. I filtered out wind speeds of zero and grouped wind speed into a three categories (1-25 km/h, 26-50 km/h and 51-75 km/h):

```{r  wind_data, eval=FALSE}

stn <- stations_search(name = "Esquimalt", interval = "hour") %>% 
  select(station_id) %>% 
  pull()
data <- weather_dl(station_id = stn)
wind_dir <- data %>% 
  select(station_id, station_name, date, wind_spd, wind_dir) %>% 
  filter(wind_spd != 0) %>% 
  mutate(wind_spd_cat = cut(wind_spd, breaks = c(0,25,50,75),
                            labels = c("1-25 km/h Winds",
                                       "26-50 km/h Winds",
                                       "51-75 km/h Winds")))
wind_dir

write_feather(wind_dir, "wind_dir.feather")
```


read from the saved feather file

```{r}
wind_dir <- read_feather("wind_dir.feather")
```


Stephanie's orginal plot:

```{r wind_plot, fig.width = 9}
wind_plot <- wind_dir %>% 
  filter(wind_dir != 0) %>% 
  mutate(wind_dir_deg = wind_dir*10) %>% 
  ggplot(aes(x = date, y = wind_dir_deg)) +
  geom_point(colour = "#3182bd", alpha = .7) +
  facet_wrap(~wind_spd_cat) +
  theme_minimal() +
  labs(x = NULL, y = "wind direction (degrees true)",
       title = "Wind Direction & Average Hourly Wind Speed Measured\nat Esquimalt Harbour (1994-2019)",
       caption = "*Data sourced from Environment Canada using the weathercan R package") +
  scale_y_continuous(limits = c(-5, 365),
                     breaks = seq(0, 360, 60),
                     expand = c(0, 0)) +
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        plot.margin = unit(c(4,50,4,4),"mm"))
wind_plot
```


### Wind speed by wind direction

Rather than date, use speed as the x axis variable

```{r}

wind_plot2 <- wind_dir %>% 
  filter(wind_dir != 0) %>% 
  mutate(wind_dir_deg = wind_dir*10) %>% 
  ggplot(aes(x = wind_dir_deg, y = wind_spd)) +
  geom_point(colour = "#3182bd", alpha = .5, position = "jitter") 
  

wind_plot2

```


Now: plot that using polar coordinates

```{r}

wind_plot2 + coord_polar()
  


```

tweak the formatting a bit

```{r}

wind_plot2 + 
  coord_polar() +
  theme_minimal() +
  scale_x_continuous(breaks = c(0, 90, 180, 270))
  


```

### Density plot

The above version is rather cluttered -- how would a density plot look?

```{r}

wind_plot2 <- wind_dir %>% 
  filter(wind_dir != 0,
         wind_spd > 0) %>% 
  mutate(wind_dir_deg = wind_dir*10) %>% 
  ggplot(aes(x = wind_dir_deg, y = wind_spd)) +
  stat_density_2d(aes(fill = stat(level)), geom = "polygon")

wind_plot2


```



```{r}

wind_plot2 + coord_polar()


```


`r Sys.Date()`