---
title: "R Notebook"
output: html_notebook
---

NOTE:

rethinking package github
https://github.com/rmcelreath/rethinking



## Attendance Data

http://www-math.bgsu.edu/~albert/jsm2018/course_files/Attendance.R


```{r}

## ------------------------------------------------------------------------
library(readr)
d <- read_csv("tribe2016.csv")


```


```{r}


## ------------------------------------------------------------------------
library(betareg)
fit <- betareg(fraction ~ Weekend + Period, data=d,
               link="logit")

## ------------------------------------------------------------------------
summary(fit)

```


### Using STAN in `rstanarm` package

What priors are used?



```{r}

## ------------------------------------------------------------------------
library(rstanarm)
fit2 <- stan_betareg(fraction ~ Weekend + Period, data=d,
                     link="logit")

```


summary of the priors

```{r}


## ------------------------------------------------------------------------
prior_summary(fit2)

```


MCMC diagnostics -- trace plots

```{r}

## ------------------------------------------------------------------------
library(bayesplot)
mcmc_trace(as.matrix(fit2))

```


```{r}

## ------------------------------------------------------------------------
mcmc_acf(as.matrix(fit2))

## ------------------------------------------------------------------------
mcmc_dens(as.matrix(fit2))


```


```{r}

## ------------------------------------------------------------------------
posterior_interval(fit2)

```

this model isn't a good fit--baseball attendance isn't uni-modal


```{r}

## ------------------------------------------------------------------------
posterior_sims <- as.matrix(fit2)
head(posterior_sims)

```




Interested in expected attendance, weekdays, each period


```{r}

## ------------------------------------------------------------------------
library(arm)
d1 <- data.frame(Label="Weekday Period 1", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"]))
d2 <- data.frame(Label="Weekday Period 2", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"] +
                      posterior_sims[, "PeriodSecond"]))
d3 <- data.frame(Label="Weekday Period 3", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"] +
                      posterior_sims[, "PeriodThird"]))

## ---- fig.height=4-------------------------------------------------------
library(ggplot2)
ggplot(rbind(d1, d2, d3), aes(Fraction)) +
  geom_density() + facet_wrap(~ Label, ncol=1)

```


shinystan ... not sure how this works

```{r}

launch_shinystan(fit2)


```

more diagnostics

- by various diagnostic measures

-- density plot doesn't fit models 

-- median, min show variation between, but sd is rather close 

- need to look at many measures to determine if it's a good model or not



```{r}

## ---- fig.height=3-------------------------------------------------------
pp_check(fit2)

## ------------------------------------------------------------------------
ynew <- posterior_predict(fit2)

## ---- fig.height=3-------------------------------------------------------
ppc_stat(d$fraction, ynew, stat="median")

## ---- fig.height=3-------------------------------------------------------
ppc_stat(d$fraction, ynew, stat="sd")

## ---- fig.height=3-------------------------------------------------------
ppc_stat(d$fraction, ynew, stat="min")


```




"Need to interact with your model...need to establish a general strategy, plan of attack, and iterate"

Would we get a different answer next time?

"Yes" >> which is why you need solid well-documented scripts! Reproducible and replicable work

If you trying to understand what drives attendance...perhaps extract season ticket holders from rest to understand variability





