---
title: "HomeRun.R"
output: html_notebook
---

## Multi-level models


### 1. Home Runs

2017 season, on May 13 -- want to predict the remainder of the season (for players who had recorded minimum of 10 PAs)

```{r}


## ------------------------------------------------------------------------
library(tidyverse)
d2017 <- filter(read_csv("may13.csv"), PA >= 50)

## ------------------------------------------------------------------------
ggplot(d2017, aes(PA, HR / PA)) + geom_point()

```

discreteness of data (integers!) of PAs on X axis, home runs on Y (i.e. bottom row is 0 home runs, next is 1 home run, then 2, etc)



> (digression: Efron and Morriss data set)
>
> https://baseballwithr.wordpress.com/2016/02/15/revisiting-efron-and-morriss-baseball-study/
>
> http://www.swarthmore.edu/NatSci/peverso1/Sports%20Data/JamesSteinData/Efron-Morris%20Baseball/EfronMorrisBB.txt
>
> https://gist.github.com/bayesball/15557fa4648ed4755500
> 
> https://vincentarelbundock.github.io/Rdatasets/doc/pscl/EfronMorris.html



#### fit the No-Pool model

```{r}

## ------------------------------------------------------------------------
fit_individual <- glm(HR ~ Player + offset(log(PA)),
                      family=poisson,
                      data=d2017)
d2017$rate_estimates <- exp(fit_individual$linear.predictors) /
                      d2017$PA

```


#### fit the Pooled Model using glm

the other extreme

```{r}


## ------------------------------------------------------------------------
fit_pool <- glm(HR ~ 1 + offset(log(PA)),
                family=poisson,
                data=d2017)
exp(coef(fit_pool)) 

```


#### Quick fit of Partial Pooling Model

see Gelman & Hill

```{r}
## ------------------------------------------------------------------------
library(lme4)
fit_pp <- glmer(HR ~ (1 | Player) + 
              offset(log(PA)),
                family=poisson,
                data=d2017)

## ------------------------------------------------------------------------
fit_pp

```


```{r}
## ------------------------------------------------------------------------
library(rstanarm)
fit_partialpool <- stan_glmer(HR ~ (1 | Player) + 
                                offset(log(PA)),
                              family=poisson,
                              data=d2017)


```




```{r}

## ------------------------------------------------------------------------
prior_summary(fit_partialpool)

```


classical method would have given a point estimate, but this approach gives a reasonable estimate range...plotted below


```{r}
## ---- fig.height=4-------------------------------------------------------
posterior <- as.matrix(fit_partialpool)
ggplot(data.frame(Sigma=sqrt(posterior[, 325])), 
       aes(Sigma)) +
      geom_density()

```




```{r}
## ---- echo=FALSE---------------------------------------------------------
shift_draws <- function(draws) {
  exp(sweep(draws[, -c(1, 325)], MARGIN = 1, 
        STATS = draws[, 1], 
        FUN = "+"))
}
Sim_Rates <- shift_draws(as.matrix(fit_partialpool))
Post_medians <- apply(Sim_Rates, 2, median)
d_estimates <- data.frame(PA = exp(fit_partialpool$offset),
                          Estimate = Post_medians,
                          Type="Partial Pool")
Estimates <- data.frame(PA = d2017$PA,
                        Estimate = d2017$rate_estimates,
                        Type="Individual")

```




```{r}
## ---- echo=FALSE---------------------------------------------------------
ggplot(rbind(d_estimates, Estimates), 
       aes(PA, Estimate, color=Type)) + geom_point() +
  geom_smooth(se=FALSE)

```




```{r}
## ------------------------------------------------------------------------
library(lme4)
library(rstanarm)
fit <- glmer(cbind(Hits, AB - Hits) ~ (1 | Player),
      family=binomial, data=bball1970)



```




