---
title: "Confidence Interval - WES reporting"
author: "Martin Monkman"
date: "February 2, 2016"
output: word_document
---

The current reporting criterial for WES is currently too arbitrary, and needs to incorporate a sliding scale. A work unit of 20 and 10 respondents will get a report, but one with 21 with 10 respondents would not.

The minimum of 10 respondents gives us assurance that confidentiality is protected. But we also need to be mindful of the _quality_ of the results; that is to say, how confident can we be in the results?

### One-off calculation

This script calculates the 95% (alpha = 0.05) confidence intervals by the work unit size, number of responses, and proportion of respondents giving a specific response.

```{r, echo = FALSE}

# set work unit size
N <- 20

# set number of respondents
n <- 10

# set the response rate [enter as decimal from 0.0 to 1.0]
RR <- 0.5

# set proportion of respondents giving specific response 
# (50% [enter as 0.5] gives the largest possible confidence interval)
P <- 0.5

# calculate the standard deviation of the sample
SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
SD

CInt <- SD * 1.96
CInt

```

The confidence interval is `r CInt`. 

### Functions

Let`s set it up as a pair of functions, one by `n` and the other by response rate (`RR`):


```{r}

# by n
ConfInt_n <- function(N, n, P){
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
}

# by response rate
ConfInt_RR <- function(N, RR, P){
  n <- N * RR
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
}

```

Now run a batch:

```{r}

for (N in 20:49){
  print(N)
}

N <- 47
RR <- 0.5
n <- trunc(N * RR)
P <- 0.5
CI <- ConfInt_RR(N, RR, P)


for(N in 20:49) {
  n <- trunc(N * RR)
  CI <- ConfInt_n(N, n, P)
  print(CI)
  }

```

### Make a table showing `N`, `n`, and the resulting confidence interval at 50% response rate.

This uses the package `dplyr`. 

#### Driven by response rate

Calculate both the `n` (based on the defined response rate) and the confidence interval (`CI`).

```{r}

library(dplyr)

N <- c(20:49)
CItable <- data.frame(N)

P <- 0.5
RR <- 0.5

mutate(CItable,
       n = trunc(N * RR),
       CI = ConfInt_n(N, n, P))


# assign the output to a table

CItable_RR <- mutate(CItable,
                  n = trunc(N * RR),
                  CI = ConfInt_n(N, n, P))

CItable_RR

```

#### Driven by response rate

Calculate the response rate `RR` (based on the number of responses) and the confidence interval (`CI`).


```{r}

P <- 0.5
n <- 10

# assign the output to a table

CItable_n <- mutate(CItable,
                    n = n,
                    CI = ConfInt_n(N, n, P))

CItable_n

```

Wait a minute -- we can make a single table with both sets of values.

```{r}

N <- c(20:49)
CItable_master <- data.frame(N)

P <- 0.5
RR <- 0.5

# assign the output to a table

CItable_master <- mutate(CItable_master,
                  n1 = trunc(N * RR),
                  CI1 = ConfInt_n(N, n1, P))

CItable_master

# add by n = 10

n2 <- 10

# assign the output to a table

CItable_master <- mutate(CItable_master,
                    n2 = n2,
                    CI2 = ConfInt_n(N, n2, P))

CItable_master


```

### Plot: N by CI

```{r}

library(ggplot2)

ggplot(CItable_master, aes(N)) + 
  geom_line(aes(y = CI1), colour = "red") + 
  geom_line(aes(y = CI2), colour = "black")

```
The end (for now)

-30-
