---
title: "Sample Size calculator"
author: "Martin Monkman"
date: "2017-08-22"
output: word_document
---


# Sample Size Calculator


This script calculates 

* CInt = confidence interval based on

* N = population size

* n = the sample size (or the number of responses, assuming 100% response)

* with P = prevalance (proportion observed in the population).



### Functions

Let's set it up as a pair of functions, one by `n` and the other by response rate (`RR`):


```{r}

# by n
ConfInt_n <- function(N, n, P){
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
  return(CInt)
}

# by response rate
ConfInt_RR <- function(N, RR, P){
  n <- n * RR
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
  return(CInt)
}




# coefficient of variation
CoefVar <- function(N, n, P){
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CoV <- (SD / P * 100)
  return(CoV)
}


```

this is the text


```{r}

N = 3000
n = 385
RR = 0.5
P = 0.1

ConfInt_n(N, n, P)

CoefVar(N, n, P)

ConfInt_RR(N, RR, P)


```



## Sample Size Calculation (based on P and CV)


```{r}

# check my equations

N <- 3000
n <- 385
P <- 0.15

# SD from proportion and sample size
SD <- sqrt((P * (1 - P)) / n)
SD
CoV <- (SD / P * 100)
CoV

# use calculated CoV and known P to calculate SD
SD <- (CoV * P / 100)
SD

# create new variable - square of SD
SD_sq <- SD^2
SD_sq 

SD_sq <- ((P * (1 - P)) / n)
SD_sq

n <- (P * (1 - P)) / SD_sq
n

n <- (P * (1 - P)) / (SD ^2)
n 

# so to simplify, should approximate n = 385:
CoV <- 12
P <- 0.15

SD <- (CoV * P / 100)
n <- (P * (1 - P)) / (SD ^2)
n 

#### ----

# ditto, with population correction

N = 3000
n = 385
P = 0.15

SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
SD
CoV <- (SD / P * 100)
CoV


# use calculated CoV and known P to calculate SD
SD <- (CoV * P / 100)
SD

# create new variable - square of SD
SD_sq <- SD^2
SD_sq 

SD_sq <- (sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))) ^ 2
SD_sq

n <- N / (((SD_sq * (N-1)) / (P * (1 - P)) + 1))
n

# so to simplify, should approximate n = 385:
N <- 3000
CoV <- 11.3
P <- 0.15

SD <- (CoV * P / 100)
n <- N / (((SD ^ 2) * (N-1)) / (P * (1 - P)) + 1)
n 

```


the SampleSizeR function

```{r}

# coefficient of variation
SampleSizeR <- function(N, P, CoV){
  SD <- (CoV * P / 100)
  n <- N / (((SD ^ 2) * (N-1)) / (P * (1 - P)) + 1)
  return(n)
}
  


```



```{r}

N <- 22500
P <- 0.05
CoV <- 10


SampleSizeR(N, P, CoV)


```



Now for the tests:


```{r, echo = FALSE}

# set population size
N <- 3000

# set sample size
n <- 385

# set the response rate [enter as decimal from 0.0 to 1.0]
RR <- 0.5

# set proportion of respondents giving specific response 
# (50% [enter as 0.5] gives the largest possible confidence interval)
P <- 0.1


```


a coefficient of variation of either 10.0% or 16.5%, 


Now run a batch:

```{r}

for (N in 20:49){
  print(N)
}

N <- 47
RR <- 0.5
n <- trunc(N * RR)
P <- 0.5
CI <- ConfInt_RR(N, RR, P)


for(N in 20:49) {
  n <- trunc(N * RR)
  CI <- ConfInt_n(N, n, P)
  print(CI)
  }

```

### Make a table showing `N`, `n`, and the resulting confidence interval at 50% response rate.

This uses the package `dplyr`. 

#### Driven by response rate

Calculate both the `n` (based on the defined response rate) and the confidence interval (`CI`).

```{r}

library(dplyr)

N <- c(20:49)
CItable <- data.frame(N)

P <- 0.5
RR <- 0.5

mutate(CItable,
       n = trunc(N * RR),
       CI = ConfInt_n(N, n, P))


# assign the output to a table

CItable_RR <- mutate(CItable,
                  n = trunc(N * RR),
                  CI = ConfInt_n(N, n, P))

CItable_RR

```

#### Driven by response rate

Calculate the response rate `RR` (based on the number of responses) and the confidence interval (`CI`).


```{r}

P <- 0.5
n <- 10

# assign the output to a table

CItable_n <- mutate(CItable,
                    n = n,
                    CI = ConfInt_n(N, n, P))

CItable_n

```

Wait a minute -- we can make a single table with both sets of values.

```{r}

N <- c(20:49)
CItable_master <- data.frame(N)

P <- 0.5
RR <- 0.5

# assign the output to a table

CItable_master <- mutate(CItable_master,
                  n1 = trunc(N * RR),
                  CI1 = ConfInt_n(N, n1, P))

CItable_master

# add by n = 10

n2 <- 10

# assign the output to a table

CItable_master <- mutate(CItable_master,
                    n2 = n2,
                    CI2 = ConfInt_n(N, n2, P))

CItable_master


```

### Plot: N by CI

```{r}

library(ggplot2)

ggplot(CItable_master, aes(N)) + 
  geom_line(aes(y = CI1), colour = "red") + 
  geom_line(aes(y = CI2), colour = "black")

```
The end (for now)

-30-
