---
title: "sample size calculations"
output: html_notebook
---

# Sample Size Calculator


```{r}

library(tidyverse)

```

This script calculates 

* CInt = confidence interval 

* CoV = coefficient of variation

* N = population size

* n = the sample size (or the number of responses, assuming 100% response)

* P = prevalance (proportion observed in the population).



### Functions: binomial distributions (P)

Three functions:

* confidence interval where `n` is the number of measurements

* confidence interval by response rate (`RR`); that is, a proportion of `n`

* coefficient of variation, using `n`



```{r}

# by n
ConfInt_n <- function(N, n, P){
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
  return(CInt)
}

# by response rate
ConfInt_RR <- function(N, RR, P){
  n <- n * RR
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CInt <- SD * 1.96
  return(CInt)
}




# coefficient of variation
CoefVar <- function(N, n, P){
  SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
  CoV <- (SD / P * 100)
  return(CoV)
}


```

a quick and dirty test

```{r}

N = 3000
n = 385
RR = 0.5
P = 0.1

ConfInt_n(N, n, P)

CoefVar(N, n, P)

ConfInt_RR(N, RR, P)


```



## Sample Size Calculation (based on P and CV)


```{r}

# check my equations

N <- 3000
n <- 385
P <- 0.15

# SD from proportion and sample size
SD <- sqrt((P * (1 - P)) / n)
SD
CoV <- (SD / P * 100)
CoV

# use calculated CoV and known P to calculate SD
SD <- (CoV * P / 100)
SD

# create new variable - square of SD
SD_sq <- SD^2
SD_sq 

SD_sq <- ((P * (1 - P)) / n)
SD_sq

n <- (P * (1 - P)) / SD_sq
n

n <- (P * (1 - P)) / (SD ^2)
n 

# so to simplify, should approximate n = 385:
CoV <- 12
P <- 0.15

SD <- (CoV * P / 100)
n <- (P * (1 - P)) / (SD ^2)
n 

#### ----

# ditto, with population correction

N = 3000
n = 385
P = 0.15

SD <- sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))
SD
CoV <- (SD / P * 100)
CoV


# use calculated CoV and known P to calculate SD
SD <- (CoV * P / 100)
SD

# create new variable - square of SD
SD_sq <- SD^2
SD_sq 

SD_sq <- (sqrt((P * (1 - P)) / n) * sqrt((N - n) / (N - 1))) ^ 2
SD_sq

n <- N / (((SD_sq * (N-1)) / (P * (1 - P)) + 1))
n

# so to simplify, should approximate n = 385:
N <- 3000
CoV <- 11.3
P <- 0.15

SD <- (CoV * P / 100)
n <- N / (((SD ^ 2) * (N-1)) / (P * (1 - P)) + 1)
n 

```


### the SampleSizeR function

Calculating `n` from desired coefficient of variation, given estimated incidence `P` and population `N`


```{r}

# coefficient of variation
SampleSizeR_CoV <- function(N, P, CoV){
  SD <- (CoV * P / 100)
  n <- N / (((SD ^ 2) * (N-1)) / (P * (1 - P)) + 1)
  return(n)
}
  

# add Confidence Interval variant


```


a quick test

```{r}

N <- 22500
P <- 0.05
CoV <- 10


SampleSizeR_CoV(N, P, CoV)


```




** batch version **

* two different values of each of `N`, `P`, and `CoV`)

(note use of `expand.grid` to build table)

```{r}


N <- c(15000, 22500)
P <- c(0.05, 0.10)
CoV <- c(10, 16.5)

variation_table <- as_tibble(expand.grid(N, P, CoV)) %>%
    rename(N = Var1,
         P = Var2,
         CoV = Var3)

variation_table <- variation_table %>%
  mutate(community_count = ifelse(N == 15000, 300, 200)) %>%
  mutate(samplesize = SampleSizeR_CoV(N, P, CoV)) %>%
  mutate(totalsample = samplesize * community_count)

variation_table


write_csv(variation_table, "variation_table.csv")

```




The end (for now)

-30-
