---
title: "SEPH test"
subtitle: "Payroll employees"
author: "Martin Monkman"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)


# tidyverse
library(tidyverse)
#library(readr)
#library(ggplot2) 
#library(dplyr)
#library(forcats)
library(stringr)
library(glue)

# monthly data series
library(lubridate)
# extending ggplot2
library("scales")

# cansim
#install.packages("devtools")
#devtools::install_github("mountainmath/cansim")
library(cansim)


### ---
# bonus function for "not in" (opposite of %in%)
# https://stackoverflow.com/a/50255505/4170032
`%!in%` <- purrr::compose(`!`, `%in%`)



```

# Introduction

Statistics Canada, Survey of Employment, Payroll^[See Statistics Canada [Survey of Employment, Payrolls and Hours](http://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=2612&dis=1)] 


Comparison with LFS^[[Comparing the Survey of employment, payrolls and hours and the Labour Force Survey](https://www150.statcan.gc.ca/n1/pub/72-203-g/2013001/part-partie8-eng.htm)]


There are multiple data tables associated with the SEPH^[[SEPH data](
https://www150.statcan.gc.ca/n1/daily-quotidien/181025/dq181025a-cansim-eng.htm)]




# Long-term trend

Up over time

B.C. has lower average weekly earning than the national average.

```{r data_read, echo = FALSE}

data_14_10_0223_01 <- get_cansim("14-10-0223-01")


# convert date field to ISO8601 with day of month (default to 01)
data_14_10_0223_01 <- data_14_10_0223_01 %>%
    mutate(REF_DATE = ymd(REF_DATE, truncated = 2)) 

data_14_10_0223_01 <- data_14_10_0223_01 %>%
  mutate(NAICS_desc =
           case_when(
             `North American Industry Classification System (NAICS)` == 
               "Health care and social assistance" ~ 
               "Health care & social assistance",
             `North American Industry Classification System (NAICS)` == 
               "Accommodation and food services" ~ 
               "Accommodation & food services",
             `North American Industry Classification System (NAICS)` == 
               "Professional, scientific and technical services" ~ 
               "Professional, scientific & technical services",
             `North American Industry Classification System (NAICS)` == 
               "Transportation and warehousing" ~ 
               "Transportation & warehousing",
             `North American Industry Classification System (NAICS)` == 
               "Administrative and support, waste management and remediation services" ~ 
               "Administrative & support",
             TRUE ~ `North American Industry Classification System (NAICS)`
           ))

# convert NAICS name to factor for ease of use later  
data_14_10_0223_01 <- data_14_10_0223_01 %>%
    mutate(NAICS_desc = 
             as_factor(NAICS_desc)) 

```


Annual trend figure in the margin (need to work on format--too many years shown)

```{r fig-margin, fig.margin = TRUE, fig.cap = "Average weekly earnings, Canada and British Columbia.", fig.width=3.5, fig.height=3.5, cache=TRUE, message=FALSE, echo = FALSE}
dataplot <- data_14_10_0223_01 %>%
  filter(GEO %in% c("British Columbia", "Canada"), 
         NAICS_desc == "Industrial aggregate excluding unclassified businesses", 
         Estimate == "Average weekly earnings including overtime for all employees") %>%
  filter(REF_DATE >= as.Date("2008-01-01")) %>%
  ggplot(aes(x=REF_DATE, y=VALUE, group = GEO, colour = GEO)) + 
    geom_line(size=1.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               minor_breaks = NULL)

dataplot2 <- dataplot +
  ylim(700, 1050) +
  scale_y_continuous(labels = comma, limits = c(700, 1050)) +
  scale_colour_manual(name=NULL,
                      breaks=c("Canada", "British Columbia"),
                      labels=c("Canada", "British Columbia"), 
                      values=c("#325A80", "#CCB550")) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour="white"),
    plot.title = element_text(face="bold"),
    legend.position=c(1,0), 
    legend.justification=c(1,0),
    legend.title = element_text(size=12),
    legend.text = element_text(size=11),
    axis.line = element_line(colour="black"),
    axis.title = element_text(size=10),
    axis.text = element_text(size=10)
  )

#SEPH_earnings_title <- as.character("Average weekly earnings including overtime for all employees")
#SEPH_earnings_subtitle <- as.character("January 2010 = 100")
SEPH_earnings_caption <- as.character("Source: Statistics Canada, CANSIM table 14-10-0223-01")
SEPH_earnings_yaxis <- as.character("Dollars")

# add titles / X-Y axis labels / caption
SEPH_earnings_plot <- dataplot2 +
  labs(
#    title = SEPH_earnings_title,
#    subtitle = SEPH_earnings_subtitle,
       caption = SEPH_earnings_caption, 
       x = NULL, y = SEPH_earnings_yaxis) 

SEPH_earnings_plot
  

```


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
