---
title: "Xaringan slide template"
subtitle: "Eigengrau theme"
author: "Martin Monkman"
date: "2018/06/06"
output:
  xaringan::moon_reader:
    css: ["default", "css_files/eigengrau.css", "css_files/eigengrau-fonts.css"]    
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: inverse, centre, middle

# A basic inverse slide.


```{r eval=FALSE, tidy=FALSE}
class: inverse, centre, middle 
```


---


# eigengrau

# h1

## h2

### h3

and some text

---


# eigengrau

This deck is based on the theme "eigengrau"

The hex code values are

- "#16161d"
- "#47474e"
- "#7e7e85"
- "#babac8"
- "#fafaff"

---
class: inverse


# This is in inverse mode

# h1

## h2

### h3

---

# This is a first level heading

## with a second level heading


???

Underneath the three question marks lie the notes ... unseen in the main slide, but visible to the presenter in presenter mode


---

An example with an image

-- need a pure black background



---

class: slide-final, left, middle

# Final slide

## Thank you

Where you might also say "questions?"


### @monkmanmh

<BR><BR>
Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan) and the eigengrau theme



